rapport résumé

### 2010-03-30 egsm.tex
an edit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@398 ccdd05d5-56c1-461a-bf98-85f6876523b4

Here we compute the bias factor $\beta$ of the sample median of a chi-square distributed variables relative to the sample geometric mean.\\ Let's define n chi-square distributed variable : $x_{l}$ l$\in\left[\hspace{-1ex}\left[\hspace{0.5ex}1,n\right]\kern-0.15em\right]$. The probability density function for each $x_{l}$ is given by : $f(x)=\frac{1}{2} e^{-x/2}$, where $$2=\langle x \rangle = \int_{0}^{\infty}xf(x)dx$$ is the \textit{expected value} of the random variable $x$. The median, $x_{1/2}$ is defined by $$ \frac{1}{2}=\int_{0}^{x_{1/2}}f(x)dx$$ which yields $x_{1/2}=2\,ln2$. Thus the bias of the \textit{expected value} is $ln2$.\\
Let's $\rho$ be the geometric sample mean of the chi-square variables : $$\rho=\left\langle(\prod_{l=1}^{n}x_{l})^{1/n}\right\rangle=\left\langle exp \left[\frac{1}{n}\sum_{l=1}^{n} \ln x_{l}\right] \right\rangle$$ so $\ln \rho$ is the arithmetic mean from the logarithm of the discret distribution and thus, for a continuous variable as $x_{l}$ we have : $$ \rho=exp \left(\int_{0}^{\infty}\ln x\,f(x)\,dx \right)$$ $$=exp \left(\frac{1}{2} \int_{0}^{\infty}e^{-x/2}\,\ln x dx \right)$$
The integral above can be calculated\footnote{\label{1}\textit{Table of integrals, series and products}, I.S. Gradshteyn/I.M. Ryzhik, BI ((256))(2).} and we find : $$ \rho\,\,=\,\frac{2}{e^{C}}= 1.122918967...\,. $$ where $C\,= \lim\limits_{s\to \infty}\left(\sum_{m=1}^{s}\frac{1}{m}-\ln s \right)$ $= 0.577215...\,.$\\
As the expected value of uncorrelated random variable is the product of the expected values (E(XY)=E(X)E(Y)), we can calculate the expected value of the geometric sample mean :$\langle \rho \rangle = \langle x_{l}^{1/n} \rangle^{n}$. The probability density function of the new random variable $Y=x^{1/n}$ is : $$P(Y)=\frac{n}{2}.e^{-Y^{n}/2}.Y^{n-1}$$ and so $$
\langle \rho \rangle = \left[\frac{n}{2}\int_{0}^{\infty}Y^{n}e^{-Y^{n}/2}dY \right]^{n}$$ $$=\frac{1}{2^{n}}\left[\int_{0}^{\infty}z^{1/n}e^{-z/2}dz\right]^{n}$$ with $z=Y^{n}$. This integral can be easily calculated\footnote{\label{2}\textit{Table of integrals, series and products},  I.S. Gradshteyn/I.M. Ryzhik, FI 2 779.} using the GAMMA function we have : $$ \langle \rho \rangle= \frac{2}{n^{n}}\left[\Gamma(\frac{1}{n})\right]^{n}$$
This result makes sense : we can calculate the exact limit as n $\rightarrow \infty$ : $\lim\limits_{n\to \infty}\langle \rho \rangle=\rho$ as illustrated on the figure below, which represent the expected value of the geometric sample mean versus n :

\vspace*{0.2cm}
\includegraphics[scale=0.19]{geosamplemean.png}

>>>>

Here we compute the bias factor $\beta$ of the sample median of a chi-square distributed variables relative to the sample geometric mean.\\ Let's define n chi-square distributed variable : $x_{l}$ l$\in\left[\hspace{-1ex}\left[\hspace{0.5ex}1,n\right]\kern-0.15em\right]$. The probability density function for each $x_{l}$ is given by : $f(x)=\frac{1}{2} e^{-x/2}$, where $$2=\langle x \rangle = \int_{0}^{\infty}xf(x)dx$$ is the \textit{expected value} of the random variable $x$. The median, $x_{1/2}$ is defined by $$ \frac{1}{2}=\int_{0}^{x_{1/2}}f(x)dx$$ which yields $x_{1/2}=2\,ln2$. Thus the bias of the \textit{expected value} is $ln2$.\\
Let's $\rho$ be the geometric sample mean of the chi-square variables : $$\rho=\left\langle(\prod_{l=1}^{n}x_{l})^{1/n}\right\rangle=\left\langle exp \left[\frac{1}{n}\sum_{l=1}^{n} \ln x_{l}\right] \right\rangle$$ so $\ln \rho$ is the arithmetic mean from the logarithm of the discret distribution and thus, for a continuous variable as $x_{l}$ we have : $$ \rho=exp \left(\int_{0}^{\infty}\ln x\,f(x)\,dx \right)$$ $$=exp \left(\frac{1}{2} \int_{0}^{\infty}e^{-x/2}\,\ln x dx \right)$$
The integral above can be calculated\footnote{\label{1}\textit{Table of integrals, series and products}, I.S. Gradshteyn/I.M. Ryzhik, BI ((256))(2).} and we find : $$ \rho\,\,=\,\frac{2}{e^{C}}= 1.122918967...\,. $$ where $C\,= \lim\limits_{s\to \infty}\left(\sum_{m=1}^{s}\frac{1}{m}-\ln s \right)$ $= 0.577215...\,.$\\
As the expected value of uncorrelated random variable is the product of the expected values (E(XY)=E(X)E(Y)), we can calculate the expected value of the geometric sample mean :$\langle \rho \rangle = \langle x_{l}^{1/n} \rangle^{n}$. The probability density function of the new random variable $Y=x^{1/n}$ is : $$P(Y)=\frac{n}{2}.e^{-Y^{n}/2}.Y^{n-1}$$ and so $$
\langle \rho \rangle = \left[\frac{n}{2}\int_{0}^{\infty}Y^{n}e^{-Y^{n}/2}dY \right]^{n}$$ $$=\frac{1}{2^{n}}\left[\int_{0}^{\infty}z^{1/n}e^{-z/2}dz\right]^{n}$$ with $z=Y^{n}$. This integral can be easily calculated\footnote{\label{2}\textit{Table of integrals, series and products},  I.S. Gradshteyn/I.M. Ryzhik, FI 2 779.} using the GAMMA function we have : $$ \langle \rho \rangle= \frac{2}{n^{n}}\left[\Gamma(\frac{1}{n})\right]^{n}$$
This result makes sense : we can calculate the exact limit as n $\rightarrow \infty$ : $\lim\limits_{n\to \infty}\langle \rho \rangle=\rho$ as illustrated on the figure below, which represent the expected value of the geometric sample mean versus n:
\resizebox{\linewidth}{!}{\includegraphics{geosamplemean.png}}
